---
title: "Data 607 Homework 1"
output:
  html_document:
    df_print: paged
---

# Introduction
This dataset is from FiveThirtyEight which is a website that collects data to use for statistical analysis.
The dataset that I choose from FiveThirtyEight is the [Polls dataset](https://github.com/fivethirtyeight/data/tree/master/polls) and specificially the approval poll ratings.
This dataset is a collection of data from different polling organization which aggregates them into this one file. 


```{r}
x = read.csv(url("https://projects.fivethirtyeight.com/polls-page/president_approval_polls.csv"))

head(x)
```

```{r}
str(x)
```

We can see that the data is not up to date as Donald Trump is the only president in this data file
```{r}
unique(x$politician)
```

```{r}
unique(x$state)
unique(x$tracking)
unique(x$sponsor_candidate)
```

# Creation of Subset
First I am going to trim down the dataset for only the columns that are needed. Some columns like pollster_rating_id, sponsor_candidate, and tracking are most likely not needed in the final subset as they do not help further the exploration of the data. Some columns like state does not have any information in them and will also be removed

```{r}
subset = x[c("politician", "pollster", "fte_grade", "sample_size", "population", "methodology", "start_date", "end_date", "created_at")]
```

Looking at the column called population there is a seperate table which tells you what each letter represents and I will be joining that into the subset in order to make it easier to understand what the codes mean
```{r}
unique(subset$population)

population_key = data.frame(c("a", "rv", "v", "lv"), c("Adults", "Registered Voters", "Voters", "Likely Voters"))
colnames(population_key) = c("population", "population_description")

subset = merge(subset, population_key, by="population")

head(subset)
```
# Conclusion
With this dataset one can graph the approval rates over time in order to see the historical approval rating of the president. One can also put specific filters based on whether they only want a certain sample size, population type, or grade of the pollster. Based on the different filters applied is can change the results drastically and which can change the public perception. I would also like to see if the dataset would be able to include the parties affiliation as that also may have a affect as maybe some pollsters only interview certain groups of people.
